---
title: "LAGOS Spatial Analysis"
author: "Matthew Ross"
date: "9/11/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
library(tidyverse) # Tidy packages
library(sf) #Spatial package that can read and create shapefiles 
library(mapview) #Interactive maps
library(LAGOSNE) #Lots and lots of clean lake data
library(USAboundaries) #USA states and counties
library(ggthemes)
```


# LAGOS Analysis


## Loading in data


### First download and then specifically grab the locus (or site lat longs)
```{r data-read}
#Lagos download script
#lagosne_get(dest_folder = LAGOSNE:::lagos_path())

#Load in lagos
#lagos <- lagosne_load()

#Grab the lake centroid info
#lake_centers <- lagos$locus

load('lake_centers.Rdata')  # work around for lagosne_get error
```



### Convert to spatial data
```{r}
#Look at the column names
#names(lake_centers)

#Look at the structure
#str(lake_centers)

#View the full dataset
#View(lake_centers %>% slice(1:100))

spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326) %>%
  st_transform(2163)

#Subset for plotting
subset_spatial <- spatial_lakes %>%
  slice(1:100)

#Dynamic mapviewer
mapview(subset_spatial)

```


### Subset to only Minnesota

```{r}
states <- us_states()

#Plot all the states to check if they loaded
#mapview(states)
minnesota <- states %>%
  filter(name == 'Minnesota') %>%
  st_transform(2163)

#Subset lakes based on spatial position
minnesota_lakes <- spatial_lakes[minnesota,]

#Plotting the first 1000 lakes
minnesota_lakes %>%
  slice(1:1000) %>%
  arrange(lake_area_ha) %>%
  mapview(.,zcol = 'lake_area_ha')
```



# In-Class work


## 1) Show a map outline of Iowa and Illinois (similar to Minnesota map upstream)

```{r }
#Plot all the states to check if they loaded
#mapview(states)
IA_IL <- states %>%
  filter(name %in% c('Iowa', 'Illinois')) %>%
  st_transform(2163)

mapview(IA_IL)
```



## 2) Subset LAGOS data to these sites, how many sites are in Illinois and Iowa
combined? How does this compare to Minnesota?

```{r}
#Subset lakes in Iowa and Illinois
iowa_illinois_lakes <- spatial_lakes[IA_IL,]

length(iowa_illinois_lakes$lagoslakeid) #total number of lakes in IA and IL
```



## 3) What is the distribution of lake size in Iowa vs. Minnesota?

- Here I want to see a histogram plot with lake size on x-axis and frequency on 
y axis (check out geom_histogram)

```{r}
iowa <- states %>%
  filter(name %in% c('Iowa')) %>%
  st_transform(2163)
  
iowa_lakes <- spatial_lakes[iowa,] %>%
  mutate(state = 'Iowa')

minnesota_lakes <- minnesota_lakes %>% 
  mutate(state = 'Minnesota')

iowa_minnesota_lakes <- rbind(iowa_lakes, minnesota_lakes)

iowa_minnesota_lakes %>%
ggplot(aes(log10(lake_area_ha))) +
  geom_histogram(bins=50) +
  facet_wrap(~state, nrow=2) +
  theme_few() +
  xlab(bquote('Lake Area ['*log[10]*'(Ha)]')) +
  ylab('Frequency')
```


## 4) Make an interactive plot of lakes in Iowa and Illinois and color them
by lake area in hectares

```{r}
#Plotting the first 1000 lakes in the Iowa and Illinois data sets
iowa_illinois_lakes %>%
  slice(1:1000) %>%
  arrange(lake_area_ha) %>%
  mapview(.,zcol = 'lake_area_ha')
```


## 5) What other data sources might we use to understand how reservoirs and 
natural lakes vary in size in these three states? 


